Shallow EDSLs and Object-Oriented Programming

### 2016-06-13 paper/paper.tex
introduction

This paper argues that OOP abstractions, including \emph{inheritance}
and \emph{subtyping}, increase the modularity and reuse of shallow
DSLs when compared to classical procedural abstraction. We make this
argument by taking a recent paper by Gibbons and Wu, where procedural
abstraction is used in Haskell to model a simple EDSL, and we recode
that EDSL in Java. Although from the \emph{syntactical} point of view
there are obvious disadvantages in the Java version, from the semantic
and modularity point of view the Java version has clear advantages.
To alleviate some of the syntactical disadvantages of Java, we create
an annotation that uses transparent code generation techniques to
automatically eliminate large portions of boilerplate code.
To further illustrate the applicability of our techniques, we conduct 
several case studies using larger DSLs from the literature.

\end{abstract}

>>>>

This paper argues that OOP abstractions, including \emph{inheritance}
and \emph{subtyping}, increase the modularity and reuse of shallow
DSLs when compared to classical procedural abstraction. We make this
argument by taking a recent paper by Gibbons and Wu, where procedural
abstraction is used in Haskell to model a simple EDSL, and we recode
that EDSL in Java. Although from the \emph{syntactical} point of view
there are obvious disadvantages in the Java version, from the semantic
and modularity point of view the Java version has clear advantages.
To alleviate some of the syntactical disadvantages of Java, we create
an annotation inspired by \emph{family polymorphism} and a recent
solution to the Expression Problem. 
The annotation uses transparent code generation techniques to
automatically eliminate large portions of boilerplate code.
To further illustrate the applicability of our tool and techniques, we conduct 
several case studies using larger DSLs from the literature. 
\end{abstract}
